\chapter{Basic maintenance}

\section{Remote power control}
Sometimes it is needed to cut the power to a part of the SALSA system,
for example to reboot the USRP. The SALSA electronics are connected to an electric outlet which
can be remotely controlled via internet at 
\verb! http://pwcontrol2.oso.chalmers.se!

\section{RIO}
RIO 47200 from Galil, the command reference is heref: 
\begin{verbatim}
http://www.galil.com/download/comref/com47xxx/
index.html#cover.html+RIO!
\end{verbatim}
To revise the RIO code, to as follows:
Run galiltools as a terminal. When connected to the RIO, press Upload to read
the current DMC code from the RIO. Then change the program as you want and
save it locally on the computer TO replace the RIO software, you first need
to turn off any ongoing loops. This can be done by running the four commands
AB0, AB1, AB2, AB3 in sequence (i.e. AB0 ENTER, then next) to abort the
four running threads. You can send the command TB to check the status
of the controller, if 127 then things are still running.
After all is aborted, you can press download to download the editor code to
the RIO. To keep it there after a restart, run BP to burn it to memory.
Finally, run RS to reset all to default power on setting.
Note that print statements in the code may cause python communication to fail
so be careful to comment those before burning the code for production use.
Output can be useful for testing though.

\section{Weather}
See http://www.oso.chalmers.se/weather/ and try to keep below 10m/s.

\section{Usage report}
To compile a usage report, go to the Usage Report page on the admin pages in the
SALSA portal. Then, enter date interval to select. Note that long interval, such as a year,
may take a few seconds to load. When loaded, open up LibreOffice Calc. 
With the pointer, mark the table on the webpage and copy it. then
Paste it in a Libreoffice Calc sheet. 
You now have a column with time for each booking. Problem is, the booking
time should be doubled if we have two telescopes for the session. One 
way of doing this is the following> add a new colum besides the one with times.
In one cell in this column, insert
\begin{verbatim}
=IF(OR(B246="Vale, Brage",B246="Brage, Vale "),A246*2,A246)
\end{verbatim}
with 246 being the row numbers, change accordingly.
Then drag downwards, and it will follow the relative cell paths,
causing all duplicated bookings to show dopuble hours. Then, sum
the new time column at the bottom, and you have a total time
for your date range.
